**The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models**

### 2021-05-09 Manuscripts/Manuscript v5/Manuscript_AM_v5.Rmd
Working on post-hoc comparisons

lm1<-lme(StO2_sim ~ Group + Day + Group:Day,
                 data = dat_sim,
                 random = ~ 1 | subject,
                 correlation = corCompSymm(form = ~ 1 | subject)
  )

@@

#creates a dataframe using the length of the covariates for the rm-ANOVA
lm_predict<-expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1),
                        subject=factor(rep(1:10)),
                          )
lm_predict$subject<-factor(paste(lm_predict$subject, lm_predict$Group, sep = "-"))

@@

lm_predict<-lm_predict%>%
    mutate(fit = predict(lm1,lm_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(lm1, lm_predict,se.fit = TRUE,type='response')$se.fit)

#plot smooths and confidence interval for GAM
f3<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),fill=Group),
              alpha=0.3,
              data=gam_predict,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=f_predict,
              show.legend = FALSE)+
  #facet_wrap(~Group)+
  labs(y=expression(paste('StO'[2],' (simulated)')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
      scale_color_aaas()+
  scale_fill_aaas()

@@

sim_linear_plot<-f4 & theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )
```

>>>>

#lm2<-lme(StO2_sim ~ Group + Day + Group:Day,
#                 data = dat_sim,
#                 random = ~ 1 | subject,
#                 correlation = corCompSymm(form = ~ 1 | subject)
#  )

@@

#creates a dataframe using the length of the covariates for the rm-ANOVA
lm_predict<-expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = c(0,2,5,7,10),
                        subject=factor(rep(1:10)),
                          )
lm_predict$subject<-factor(paste(lm_predict$subject, lm_predict$Group, sep = "-"))

@@

#using lm
lm_predict<-lm_predict%>%
    mutate(fit = predict(lm1,lm_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(lm1, lm_predict,se.fit = TRUE,type='response')$se.fit)

#using lme
#lm_predict<-lm_predict%>%
#   mutate(fit = predict(lm1,lm_predict,se.fit = TRUE,type='response',level=0),
 #         se.fit = predict(lm1, lm_predict,se.fit = TRUE,type='response',level=0))

#plot smooths and confidence interval for GAM
f3<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),fill=Group),
              alpha=0.3,
              data=gam_predict,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=gam_predict,
              show.legend = FALSE)+
  #facet_wrap(~Group)+
  labs(y=expression(paste('StO'[2],' (simulated)')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
      scale_color_aaas()+
  scale_fill_aaas()

@@

sim_linear_plot<-f4 & theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

#posthoc comparisons for the linear model
library(multcomp)

summary(glht(lm1, linfct=mcp(Group="Tukey", interaction_average=TRUE)))
